Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-06-27 tex/scam/faithful.tex
added faithful.tex as main file in aos format

\documentclass{article}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,custom_math}
\usepackage[hscale=0.7,vscale=0.8]{geometry}

\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{authblk}

\usepackage{yhmath}

\bibliographystyle{plainnat}

\title{Faithful Variable Screening for High-Dimensional Convex Regression}

%\author{
%Min Xu, Minhua Chen, John Lafferty
%}

\author[1]{Min Xu}
\author[2,3]{Minhua Chen}
\author[2,3]{John Lafferty}

\affil[1]{Machine Learning Department, Carnegie Mellon University}
\affil[2]{Department of Statistics, University of Chicago}
\affil[3]{Department of Computer Science, University of Chicago}

% ENVIRONMENTS
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheoremstyle{remark}{\topsep}{\topsep}%
     {\normalfont}% Body font
     {}           % Indent amount (empty = no indent, \parindent = para indent)
     {\bfseries}  % Thm head font
     {.}          % Punctuation after thm head
     {.5em}       % Space after thm head (\newline = linebreak)
     {\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}% Thm head spec
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{definition}{Definition}[section]

@@

% John's macros
\def\X{\mathcal{X}}
\def\comma{\unskip,~}
\def\truep{p^*}
\def\div{\|\,}
\long\def\comment#1{}
\def\reals{{\mathbb R}}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Cov{\mathop{\text{Cov}}}
\def\supp{\mathop{\text{supp}\kern.2ex}}
\def\argmin{\mathop{\text{\rm arg\,min}}}
\def\arginf{\mathop{\text{\rm arg\,inf}}}
\def\argmax{\mathop{\text{\rm arg\,max}}}
\let\tilde\widetilde
\def\csd{${}^*$}
\def\mld{${}^\dag$}
\def\dos{${}^\ddag$}
\def\W{\widetilde Y}
\def\Z{\widetilde X}
\let\hat\widehat
\let\tilde\widetilde
\def\given{{\,|\,}}
\def\ds{\displaystyle}
\def\bs{\backslash}
\def\1{{(1)}}
\def\2{{(2)}}
\def\pn{{(n)}}
\def\ip{{(i)}}
\def\Xbar{\overline{X}}
\def\except{\backslash}
\def\npn{\mathop{\textit{NPN\,}}}
\def\i{{(i)}}
\def\cE{{\mathcal{C}}}
\def\cM{{\mathcal{M}}}
\def\cF{{\mathcal{F}}}
\def\cP{{\mathcal{P}}}
\def\cG{{\mathcal{G}}}
\def\tr{\mathop{\text{tr}}}
\long\def\comment#1{}
\def\ti#1{#1}
\def\titi#1{\textit{#1}}
\def\cram{{\sc cram}}
\def\spam{{\small\sc SpAM}}
\def\diag{\mathop{\rm diag}}
\def\ones{\mathbf{1}}
\def\threebars{\mbox{$|\kern-.25ex|\kern-.25ex|$}}
\def\fatnorm#1{\threebars #1 \threebars}
\def\rank{\mathop{\rm rank}}
\def\S{\mathcal{S}}
\def\H{\mathcal{H}}
\def\K{{K}}
\def\rank{\mathop{\rm rank}}
\def\half{{1/2}}
\def\Y{\mathbb{Y}}
\def\M{\mathbb{M}}
\def\F{\mathbb{F}}
\def\pinv{{-1}}
%\def\ones{\mathds{1}}
%\def\ones{1}
\def\Res{Z}
\def\Proj{P}
\def\cN{{\mathcal N}}
\def\cT{{\mathcal H}}
\def\coloneqq{:=}
\def\mathbf#1{\mbox{\boldmath $#1$}} 
\def\bar#1{\overline{#1}}

\maketitle

\begin{abstract}
We consider the problem of estimating a convex function of several variables from noisy values of the function at a finite sample of input points. Convex function estimation is subject to the curse of dimensionality where the sample size necessary for consistent estimation increases exponentially with the dimensionality of the observed variables $p$. However, if the function is sparse, with $s << p$ relevant variables, then one could achieve consistency in the high-dimensional setting by first identifying the $s$ variables. We develop a faithful screening procedure to compute a set $S$ that contains the $s$ relevant variables. Our approach is a two-stage method that estimates a sum of $p$ one-dimensional convex functions, followed by one-dimensional concave regression fits on the residuals. The method is based on quadratic programming, and in contrast to standard sparse additive models, requires no tuning parameters for smoothness. Under appropriate assumptions, we prove that the procedure is faithful in the population setting, yielding no false negatives, and we give a finite sample statistical analysis. In addition, we introduce algorithms for efficiently carrying out the required quadratic programs. The approach leads to significant computational and statistical advantages over fitting a full model, and provides an effective, practical approach to variable screening in convex regression.

\end{abstract}
\vskip10pt

@@

%\bibliographystyle{plain}
\bibliography{local}

@@

>>>>

\documentclass[12pt,pdftex,aos]{imsart}

\RequirePackage[OT1]{fontenc}
\usepackage{amsthm,amsmath,amsfonts,natbib,mathtools,custom_math}
\RequirePackage{hypernat}
\usepackage[ruled,section]{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{graphicx}
\usepackage[hscale=0.7,vscale=0.8]{geometry}

\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{yhmath}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\trace}{trace}
\DeclareMathOperator{\pr}{Prob}
\DeclareMathOperator{\cov}{Cov}
\newcommand{\EE}[1]{\mathbb{E}({#1})}
\def\interleave{|\kern-.25ex|\kern-.25ex|}
\def\interleavesub{|\kern-.15ex|\kern-.15ex|}
\newcommand{\nnorm}[1]{\interleave {#1} \interleave}
\newcommand{\nnormsub}[1]{\interleavesub {#1} \interleavesub}
\newcommand{\nNorm}[1]{\left|\kern-.25ex\left|\kern-.25ex\left| {#1}\right|\kern-.25ex\right|\kern-.25ex\right|}
\newcommand{\norm}[1]{\left\|{#1}\right\|}
%\newcommand{\trnorm}[1]{\norm{{#1}}_{\mathrm{tr}}}
\newcommand{\trnorm}[1]{\norm{{#1}}_{*}}
\newcommand{\spnorm}[1]{\norm{{#1}}_{\mathrm{sp}}}
\newcommand{\frnorm}[1]{\norm{{#1}}_{\mathrm{F}}}
%\newcommand{\trnnorm}[1]{\nnorm{{#1}}_{\mathrm{tr}}}
\newcommand{\trnnorm}[1]{\nnorm{{#1}}_{*}}
\newcommand{\spnnorm}[1]{\nnorm{{#1}}_{\mathrm{sp}}}
\newcommand{\spnnormsub}[1]{\nnormsub{{#1}}_{\mathrm{sp}}}
\newcommand{\frnnorm}[1]{\nnorm{{#1}}_{\mathrm{F}}}
\newcommand{\Norm}[1]{\left\|{#1}\right\|}
%\newcommand{\Trnorm}[1]{\Norm{{#1}}_{\mathrm{tr}}}
\newcommand{\Trnorm}[1]{\Norm{{#1}}_{*}}
\newcommand{\Spnorm}[1]{\Norm{{#1}}_{\mathrm{sp}}}
\newcommand{\Frnorm}[1]{\Norm{{#1}}_{\mathrm{F}}}
%\newcommand{\Trnnorm}[1]{\nNorm{{#1}}_{\mathrm{tr}}}
\newcommand{\Trnnorm}[1]{\nNorm{{#1}}_{*}}
\newcommand{\Spnnorm}[1]{\nNorm{{#1}}_{\mathrm{sp}}}
\newcommand{\Frnnorm}[1]{\nNorm{{#1}}_{\mathrm{F}}}
\newcommand{\colspan}[1]{\mathrm{colspan}\left({#1}\right)}
\def\myllangle{\langle\mspace{-4mu}\langle\mspace{1mu}}
\def\myrrangle{\mspace{1mu}\rangle\mspace{-4mu}\rangle}

\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheoremstyle{remark}{\topsep}{\topsep}%
     {\normalfont}% Body font
     {}           % Indent amount (empty = no indent, \parindent = para indent)
     {\bfseries}  % Thm head font
     {.}          % Punctuation after thm head
     {.5em}       % Space after thm head (\newline = linebreak)
     {\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}% Thm head spec
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{definition}{Definition}[section]

\renewcommand{\baselinestretch}{1.1}
%\parskip12pt
%\parindent0pt
\setcounter{tocdepth}{2}

% John's macros
\def\X{\mathcal{X}}
\def\comma{\unskip,~}
\def\truep{p^*}
\def\div{\|\,}
\long\def\comment#1{}
\def\reals{{\mathbb R}}
\def\R{\reals}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Cov{\mathop{\text{Cov}}}
\def\supp{\mathop{\text{supp}\kern.2ex}}
\def\argmin{\mathop{\text{\rm arg\,min}}}
\def\arginf{\mathop{\text{\rm arg\,inf}}}
\def\argmax{\mathop{\text{\rm arg\,max}}}
\let\tilde\widetilde
\def\csd{${}^*$}
\def\mld{${}^\dag$}
\def\dos{${}^\ddag$}
\def\W{\widetilde Y}
\def\Z{\widetilde X}
\let\hat\widehat
\let\tilde\widetilde
\def\given{{\,|\,}}
\def\ds{\displaystyle}
\def\bs{\backslash}
\def\1{{(1)}}
\def\2{{(2)}}
\def\pn{{(n)}}
\def\ip{{(i)}}
\def\Xbar{\overline{X}}
\def\except{\backslash}
\def\npn{\mathop{\textit{NPN\,}}}
\def\i{{(i)}}
\def\cE{{\mathcal{C}}}
\def\cM{{\mathcal{M}}}
\def\cF{{\mathcal{F}}}
\def\cP{{\mathcal{P}}}
\def\cG{{\mathcal{G}}}
\def\tr{\mathop{\text{tr}}}
\long\def\comment#1{}
\def\ti#1{#1}
\def\titi#1{\textit{#1}}
\def\cram{{\sc cram}}
\def\spam{{\small\sc SpAM}}
\def\diag{\mathop{\rm diag}}
\def\ones{\mathbf{1}}
\def\threebars{\mbox{$|\kern-.25ex|\kern-.25ex|$}}
\def\fatnorm#1{\threebars #1 \threebars}
\def\rank{\mathop{\rm rank}}
\def\S{\mathcal{S}}
\def\H{\mathcal{H}}
\def\K{{K}}
\def\rank{\mathop{\rm rank}}
\def\half{{1/2}}
\def\Y{\mathbb{Y}}
\def\M{\mathbb{M}}
\def\F{\mathbb{F}}
\def\pinv{{-1}}
%\def\ones{\mathds{1}}
%\def\ones{1}
\def\Res{Z}
\def\Proj{P}
\def\cN{{\mathcal N}}
\def\cT{{\mathcal H}}
\def\coloneqq{:=}
\def\mathbf#1{\mbox{\boldmath $#1$}} 
\def\bar#1{\overline{#1}}

@@

\begin{frontmatter}
\title{Faithful Variable Screening for High-Dimensional Convex Regression}
\runtitle{Faithful Variable Screening for Convex Regression}

\begin{aug}
\vskip10pt
\author{\fnms{Min} \snm{Xu}\ead[label=e1]{minx@cs.cmu.edu}}
\comma 
\author{\fnms{Minhua} \snm{Chen}\ead[label=e2]{minhua@galton.uchicago.edu}}
\and
\author{\fnms{John} \snm{Lafferty}\ead[label=e3]{lafferty@galton.uchicago.edu}}
\address{
\begin{tabular}{ccc}
\\[10pt]
Machine Learning Department & Department of Statistics & Department of Statistics\\
Carnegie Mellon University  & University of Chicago & Department of Computer Science\\
& & University of Chicago
\end{tabular}
\\[20pt]
\printead{e1,e2,e3}}
\end{aug}

\begin{abstract}
  We consider the problem of variable selection in convex nonparametric
  regression.  Under the assumption that the true regression
  function is sparse and convex, we
  develop a faithful screening procedure to compute a set $S$ that
  contains the $s$ relevant variables. Our approach is a two-stage
  method that estimates a sum of $p$ one-dimensional convex functions,
  followed by one-dimensional concave regression fits on the
  residuals. The method is based on quadratic programming, and in
  contrast to standard sparse additive models, requires no tuning
  parameters for smoothness. Under appropriate assumptions, we prove
  that the procedure is faithful in the population setting, yielding
  no false negatives, and we give a finite sample statistical
  analysis. In addition, we introduce algorithms for efficiently
  carrying out the required quadratic programs. The approach leads to
  significant computational and statistical advantages over fitting a
  full model, and provides an effective, practical approach to
  variable screening in convex regression.
\end{abstract}

\begin{keyword}
\kwd{nonparametric regression}
\kwd{convex regression}
\kwd{variable selection}
\kwd{quadratic programming}
\kwd{additive model}
\end{keyword}

%\tableofcontents
\vskip20pt
\end{frontmatter}

\maketitle
\vskip10pt

@@

@@

\section*{Acknowledgements}
Research supported in part by NSF grants IIS-1116730, 
AFOSR grant FA9550-09-1-0373, ONR grant
N000141210762, and an Amazon AWS in Education Machine Learning
Research grant.

\bibliographystyle{imsart-nameyear}
\bibliography{local}
